---
title: "Obtaining 489 cell type gene signatures from Aran et al., 2017"
author:
- name: Kevin Rue-Albrecht
  affiliation: 
  - &id4 Kennedy Institute of Rheumatology, University of Oxford,
    Headington, Oxford OX3 7FY, UK.
  email: kevin.rue-albrecht@kennedy.ox.ac.uk
date: "`r BiocStyle::doc_date()`"
package: "`r BiocStyle::pkg_ver('xCellData')`"
bibliography: bibliography.bib
output: 
  BiocStyle::html_document:
    toc_float: true
abstract: |
  Instructions on how to obtain 489 cell type gene signatures from Aran et al., 2017.
vignette: |
  %\VignetteIndexEntry{Obtaining 489 cell type gene signatures from Aran et al., 2017}
  %\VignetteEngine{knitr::rmarkdown}
---

```{r, echo=FALSE, results="hide", message=FALSE}
require(knitr)
opts_chunk$set(error=FALSE, message=FALSE, warning=FALSE)
```

```{r style, echo=FALSE, results='asis'}
BiocStyle::markdown()
```

# Introduction 

The `r Githubpkg("kevinrue/xCellData")` package provides a _R_ /
_Bioconductor_ resource for obtaining and representing 489 cell type gene signatures from [@Aran2017].

This packages uses the `r Githubpkg("kevinrue/unisets")` `BaseSets` class to represent the collection of signatures.
However, the data itself is distributed with the package as a GMT file, which may be parsed and imported by other packages (e.g. `r Biocpkg("GSEABase")` `GeneSetCollection`, `r Githubpkg("Kayla-Morrell/GeneSet")` `tbl_geneset`).

# Data preprocessing

The script used to download and preprocess the data is distributed with the package.
You can find it at the following location:

```{r}
system.file(package = "xCellData", "scripts", "makeData.R")
```

Briefly, the script downloads "Additional file 3: The 489 cell type gene signatures. (XLSX 417 kb)" from the [https://genomebiology.biomedcentral.com][genomebiology_aran_add3] website and reformats the content of the published Microsoft Excel file into a GMT text file.

# Workflow

## Loading the data

We use the `xCellData()` function to parse the GMT file distributed with the package into a `r Githubpkg("kevinrue/unisets")` `BaseSets` object.

```{r}
library(xCellData)
library(unisets)
xsig <- xCellData()
xsig
```

## Using the data

The signatures may then be used for downstream analyses such as cell type annotation.

For instance, the `BaseSets` object can be split into a list of signatures, for use in functions such as `lapply`.

```{r}
as.list(xsig)
```

One may also inspect the number of genes in each signature.

```{r}
dat <- setLengths(xsig)
hist(
    dat, breaks = 100, xlim=c(0, max(dat)),
    main = "Distribution of signature sizes", xlab = "Number of genes"
)
```

Example of packages using `r Githubpkg("kevinrue/xCellData")` include:

- `r Githubpkg("kevinrue/hancock")`

# Session information

```{r}
sessionInfo()
```

[genomebiology_aran_add3]: https://genomebiology.biomedcentral.com/articles/10.1186/s13059-017-1349-1#MOESM3

# References
